The Community Land Model, version 5.1 \\ One-at-a-time Parameter Perturbation Ensemble

### 2024-08-12 ms/ms.tex
overleaf

A small fraction of parameters tend to explain a large amount of the ensemble variance (Figure \ref{fig:variance}) for any given variable or metric. For example, just four parameters (maximum leaf wetted fraction, liquid canopy storage, the turbulent transfer coefficient, and leaf characteristic length) explain upwards of 90\% of the canopy evaporation variance across all the forcing scenarios (Supp Fig S3). Conversely, GPP, which is a significantly more complicated process, requires 20 parameters to explain 90\% of the control ensemble variance. In general, most of the parameter perturbations had a relatively small impact on any given output variable. Though this experiment encompasses 211 parameters, no single variable responds to all 211 parameters. In fact, the effective parameter space dimensionality for any given output variable tends to be much smaller than 211. While the principle of land models is that these are all interlinked via complex feedback processes, many parameters nonetheless have impacts that are mostly limited to their own domain.

@@

The third challenge involved managing computational cost.  Quantifying parameter effects is a necessary prerequisite to automated calibration and uncertainty quantification. Because the parameter space of CLM is quite large and the model response is potentially non-linear, large ensembles are required to adequately resolve response surfaces. With standard CLM configurations, such ensembles would far exceed our computing resources. As such, the computational cost of inferring parameter effects is a major constraint. We were able to reduce ensemble generation computational cost 500x by strategically reducing the number of model grid cells (Figure \ref{fig:sg}) and by leveraging a Semi-Analytic Spin-Up (SASU) offered by the matrix approach to land biogeochemistry models (Lu et al. 2020, Luo 2022, Liao et al. 2023) and a linearized spin-up solver (see Sections \ref{sect:mcn} and \ref{sect:sg} for details). As long as computational resources remain constrained, designing faster model configurations and more efficient sampling strategies will be important for effective model calibration and uncertainty quantification.

@@

Land only caveat? Our PPE is run as land-only simulations, using prescribed atmospheric forcing. This is likely to be adequate for estimating many land surface fluxes and pools, in particular of carbon, but could lead to biased estimates of water and energy fluxes, as we know that atmospheric feedbacks to land surface characteristics modify the state of the land surface itself (Laguë et al. 2019). Additional work is planned to leverage this one-at-a-time land-only PPE to quantify how much, where, and for which processes atmospheric feedbacks are likely to change the land surface state substantially.

@@

We have likewise begun a follow-on activity that perturbs a subset of important parameters with the Latin hypercube sampling strategy \cite{mckay2000} to work towards routine model calibration.  Our first foray into exploring the potential for global calibration is with leaf area index due to controls that this variable has on biogeophysical and biogeochemical fluxes. By investing in the infrastructure that we introduce here, all of our subsequent parameter perturbation experiments have required much less time and effort. We expect to continue to extend our efforts in this domain towards model calibration, as well as repeat the foundational one-at-a-time experiments with subsequent model releases.

>>>>

A small fraction of parameters tend to explain a large amount of the ensemble variance for any given variable or metric (Figure \ref{fig:variance}). For example, just four parameters (maximum leaf wetted fraction, liquid canopy storage, the turbulent transfer coefficient, and leaf characteristic length) explain upwards of 90\% of the canopy evaporation variance across all the forcing scenarios (Supp Fig S3). Conversely, GPP, which is a significantly more complicated process, requires 20 parameters to explain 90\% of the control ensemble variance. In general, most of the parameter perturbations had a relatively small impact on any given output variable. Though this experiment encompasses 211 parameters, no single variable responds to all 211 parameters. In fact, the effective parameter space dimensionality for any given output variable tends to be much smaller than 211. While the principle of land models is that these are all interlinked via complex feedback processes, many parameters nonetheless have impacts that are mostly limited to their own domain.

@@

The third challenge involved managing computational cost.  Quantifying parameter effects is a necessary prerequisite to automated calibration and uncertainty quantification. Because the parameter space of CLM is quite large and the model response is potentially non-linear, large ensembles are required to adequately resolve response surfaces. With standard CLM configurations, such ensembles would far exceed our computing resources. As such, the computational cost of inferring parameter effects is a major constraint. We were able to reduce ensemble generation computational cost 500x by strategically reducing the number of model grid cells (Figure \ref{fig:sg}) and by leveraging a Semi-Analytic Spin-Up (SASU) offered by the matrix approach to land biogeochemistry models (Lu et al. 2020, Luo 2022, Liao et al. 2023) and a linearized spin-up solver (see Sections \ref{sect:mcn} and \ref{sect:sg} for details). As long as computational resources remain constrained, designing faster model configurations and efficient sampling strategies will be important for effective model calibration and uncertainty quantification.

@@

Our PPE is run as land-only simulations, using prescribed atmospheric forcing. This is likely to be adequate for estimating many land surface fluxes and pools, in particular of carbon, but could lead to biased estimates of water and energy fluxes, as we know that atmospheric feedbacks to land surface characteristics modify the state of the land surface itself (Laguë et al. 2019). Additional work is ongoing to extend this work in a coupled model framework to quantify how much, where, and for which processes atmospheric feedbacks are likely to modulate the land parameter effects substantially.

@@

We have likewise begun a follow-on activity that perturbs a subset of important parameters with the Latin hypercube sampling strategy \cite{mckay2000} to work towards routine model calibration. By investing in the infrastructure that we introduce here, all of our subsequent parameter perturbation experiments have required much less time and effort. We expect to continue to extend our efforts in this domain towards model calibration and uncertainty quantification, as well as repeating the foundational one-at-a-time experiments with subsequent model releases.
